---
title: "On the estimation of COVID-19"
author: "Jialin Lu"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

`r sans_serif(
"I give a web tool to compute the estimated total number of cases, by allowing you to interactively adjust the predefined values.")
`

```{r, echo=FALSE}
htmltools::includeHTML("figure_1.html")
```


# Introduction

@Imperial2020, a research group from Imperial College London, made a simple estimation of the total number of the novel Coronavirus (*2019-nCoV*) cases stemmed recently in Wuhan City, China. The method of estimation is in general valid and sound^[I will say it is too naive and simple]. Yet it does have certain degree of concerns:

1. This model is too simple (in other words too general) anyway, a lot of features like the information of sequential dynamics is not utilized or considered at all.
1. as for most of epidemiological methods, this estimation requires many underlying asumptions and approximations, although most of which seem justified to me.

Here I provided an interactive estimation, where you can manipulate the values of assumptions and see how this will change the final result accordingly^[For example, if there are more abroad cases identified.]


## How is it computed?

Basically, the estimated total number is computed by *maximum likelihood* estimation(MLE). But since it is under a bionomial assumption, in fact it is just computed in a very straight forward manner.


```{r}
  num_intl_case = 3 # Exported number of confirmed cases
  num_intl_passenger = 3301 # Daily international passengers travelling out of Wuhan International airport
  catchment_in_million = 19 # (in million) Effective catchment population of Wuhan airport
  catchment = catchment_in_million * 10^6
  detection_window = 10 # (in days) Detection window

  p_intl_daily = num_intl_passenger / catchment # daily prob of international travel
  p = p_intl_daily * detection_window # the probability any one case will be detected overseas

  num_case = num_intl_case / p # the final result, total number of cases,
  print(num_case)
  num_case = as.integer(num_case)
  print(num_case)
```

## About the assumptions

There are, in fact, many other assumptions underlying the performed esimation.How would the number change, if the assumptions change?

* The detection window is 10 days`r margin_note("The longer the detection window, the smaller the estimated total number of cases")`, including incubation and the time for developing symptions into final detection. The incubation period of 2019-nCov -- 10 days -- is not known and has been approximated withthe estimates obtained for MERS-CoV and SARS.

* the population of Human is also approximated. `r margin_note("The longer the detection window, the smaller the estimated total number of cases")`

* Only counts the cases over of some degree of severity (requre hospitalisation)
* There will not be a case that the symptoms are developed outside Wuhan City and detected after returning to Wuhan.
* We ignore the time duration for international visitors is shorter and thus is less risky than the local citizens. But, if we count it, the total estimated number will be more
* We assume that international travel is independent of the risk of exposure to 2019nCoV or of infection status.

## Comment

I will not think it is a decently advanced modelling, this entirety of thing to me is too 20-century basic. But I nonetheless understand the difficulty here, you just do not have the auxiliary information to go sophisticated: you only got these simple data, then you only got to do it simply.
